Global land cover fraction mapping: methods, variables, challenges

### 2020-05-27 article/article.tex
Writing first draft of discussion, conclusions

% Minimalistic template for a generic article, geared towards MDPI guidelines
% Use PDFLaTeX
\documentclass[a4paper,10pt]{article}
% Packages provided by the MDPI template already
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{indentfirst}
\usepackage{booktabs} % For \toprule, etc.
\usepackage[sort&compress,sectionbib]{natbib}
\bibliographystyle{newapa}

@@

% Track changes package, remove when final; using a drop-in rather than the old version in TeX Live
% Usage: \added, \deleted, \replaced{new}{old}, \highlight, \comment
% All followed by an optional [id=<your id>, comment=<some comment>]
\usepackage[xcolor={usenames,dvipsnames}]{changes}
\definecolor{gem}{HTML}{003300}
\definechangesauthor[color=gem]{dm}
\definechangesauthor[color=NavyBlue]{jv}
\definechangesauthor[color=Purple]{nt}
\definechangesauthor[color=Mahogany]{mh}

@@

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS-LC100}{CGLS-LC100}{Copernicus Global Land Services 100 m Land Cover}
\newacronym{CGLS}{CGLS}{Copernicus Global Land Services}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{IIASA}{IIASA}{International Institute for Applied Systems Analysis}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym{OSAVI}{OSAVI}{Optimised Soil-Adjusted Vegetation Index}
\newacronym{NDVI}{NDVI}{Normalised Difference Vegetation Index}
\newacronym{NIR}{NIR}{near infra-red}
\newacronym{SCM}{SCM}{subpixel confusion-uncertainty matrix}
\newacronym{FNC}{FNC}{fuzzy nearest centroid}
\newacronym{PLS}{PLS}{partial least squares}
\newacronym{MLR}{MLR}{multinomial logistic regression}
\newacronym{RF}{RF}{random forest}
\newacronym{TOC}{TOC}{Top-of-Canopy}
\newacronym{CART}{CART}{Classification and regression tree}
\newacronym{GLM}{GLM}{General linear regression model}
\makenoidxglossaries
% Disable abbreviating RF for now [text={random forest}]
%\glsunset{RF}

@@

@@

\begin{enumerate}
 \item Relevance of fractional LC mapping (link to user requirements and SDGs)
 \item Studies that have looked into fractional LC mapping so far
 \item Available fractional methods that were introduced in literature (wavelet transform etc.)
 \item Methods that are tested in this paper
 \item Global land cover maps
\end{enumerate}

Land cover, as one of the key variables for monitoring a number of \glspl{SDG}, has received attention in recent years, and new global land cover maps have emerged to better map the current land cover of the world, as well as to track land cover change.
Some of the recent achievements have been the \ac{LC-CCI} map \highlight[id=dm, comment=Is there a better citation?]{\citep{defourny2012cci}} that aims at maintaining a long medium-resolution time series of global land cover for the climate community, \highlight[id=dm, comment=ditto]{\ac{CGLS-LC100}} map that aims at a finer spatial resolution and high quality with yearly updates since 2015, and the \ac{FROM-GLC} map \citep{fromglc2019} that showcases the potential of land cover mapping at 10 m resolution.

@@

A potential solution to this issue is land cover fraction mapping, where instead of a discrete class, the proportion of every class in the legend is reported for every pixel of the map.
This is also called ``fuzzy'' or ``soft'' classification, and sometimes ``subpixel'' classification or ``linear mixture modelling'' \citep{Okeke2006fuzzyexponent}. A related term is ``super
Land cover fraction mapping has been attempted in the past, both in research \citep{adams_classification_1995, foody1996fuzzyevaluation, colditz_land_2011, sharma_assessing_2011, uma_shankar_wavelet-fuzzy_2011, dwivedi_optimisation_2012, lizarazo_quantitative_2012, gessner_estimating_2013, okujeni_generalizing_2018} and production (e.g. \citealp{hansen2000hardtree, hansen_continuous_2011, pengra_global_2015, hansen_global_2003, sexton_global_2013}).
Nevertheless, to date only \ac{CGLS-LC100} provides a global map with fractions of every major land cover class.

@@

%The second drawback of current global land cover products is that their classification accuracy tends to be low, averaging at around 65\% \citep{tsendbazar2016integrating}. The third drawback is that all of these products use what is known as ``hard'' or ``crisp'' classification: each pixel is assigned to one particular land cover class only. The result of a hard classification is a thematic map. This classification type treats each pixel as homogeneous with regards to land cover, which is rarely the case in reality. In contrast, a ``fuzzy'' or ``soft'' classification, sometimes also called ``subpixel'' classification or ``linear mixture modelling'' \citep{Okeke2006fuzzyexponent} identifies the proportion of each land cover class within each pixel. The result of a fuzzy classification is one map per land cover class, showing the proportion of the pixel area covered by the land cover class.

%Hard classification is not well-suited for coarse resolution imagery, due to a large proportion of mixed pixels in it (representing areas of mixed land cover), compared to endmember pixels (or pure pixels, which are dominated by a single land cover class). When hard classification is attempted on such imagery, the accuracy of the result can be no higher than the area fraction that the dominant land cover type of the pixel occupies \citep{latifovic2004accuracy}. There have been attempts at increasing accuracy by defining mosaic classes, such as mixed forests (50\% needleleaf and 50\% broadleaf trees), but it leads to a proliferation of classes that increases map complexity and still does not deal with the core problem, thus not improving the classification accuracy by much \citep{tsendbazar2016comparative}. In contrast, ``fuzzy'' or ``soft'' classification results in each pixel containing information about the proportion of each class within that pixel, therefore operating on sub-pixel scales. This has the advantage of representing land cover more accurately, and gives the ability to represent mosaic classes as a combination of pure classes instead. Since the output of fuzzy classification is one raster per land cover class, with smooth transitions between classes, it is suitable for more in-depth analysis and user-specific visualisation criteria \citep{tsendbazar2016integrating}. Despite that, hard classification is still the most often used classification type due to the number of algorithms developed for it, ease of storing the result (it is thematic and thus takes little storage space), displaying it in a single map (albeit in a less accurate fashion) and performing accuracy assessment.

%There is a number of different classification algorithms, but only several of them are suitable to be used for fuzzy classification \citep{nath2014methods}. The two methods most commonly used in scientific literature are fuzzy \textit{c}-means and neural networks \citep{zhang2001fullyfuzzy}. Neural networks in particular are well-suited for fuzzy classification, since they allow multiple continuous output as well as input variables in a single model \citep{foody1997fuzzynnet}. Fuzzy \textit{c}-means, also known as fuzzy \textit{k}-means or soft \textit{k}-means, is a statistical method that relies on the proximity of pixels in feature space to class centroids, and thus is also suitable for fully fuzzy classification. However, since \textit{k}-means is an unsupervised classification algorithm, the ability to make use of training data in fuzzy \textit{c}-means is limited to determining class centroids with more precision \citep{hengl2004fuzzycmeans}.

%In addition, any other algorithms that provide a measure of uncertainty about class membership can be used, such as the ratio of individual tree votes of random forest \citep{breiman2001random} or class probabilities in gradient boosting \citep{friedman2001gradientboost}. While classification uncertainty is not a direct measure of class membership \citep{sytze2000fuzzyset}, they are nonetheless correlated. As such, it has been used by numerous authors as a proxy indicator of class membership, achieving satisfactory classification accuracy \citep{foody2002accuracy}.

%Furthermore, algorithms that can handle continuous variables, but only one response variable (such as random forest), can be used as well, by creating separate models for every class. This approach is called binary relevance \citep{karalas2016br}. Using this approach, the data needs to be post-processed after classification at a pixel level to conform to physical constraints (class membership must be between 0 and 100\% and sum up to 100\%). Random forest has been reported to give higher or equal accuracy results compared to other algorithms, like Support Vector Machines and individual decision trees, in hard classification scenarios using satellite imagery similar to that of PROBA-V \citep{duro2012algorithmcomparison}. Random forest regression was also shown to perform as well as other algorithms in fuzzy classification scenarios \citep{walton2008subpixelrf}, although it is used in much fewer studies on fuzzy classification than the other algorithms mentioned previously. Gradient boosting is an algorithm related to random forest, with the same advantages and disadvantages, but it is known to perform better in machine learning challenges \citep{chen2015higgs} and has not yet been tested on fuzzy land cover classification.

%Accuracy assessment is rather straight-forward for hard classification: typically a confusion matrix is employed for this purpose, showing how many pixels (or in some cases how much area \citep{stehman2009sampling}) in the image have been classified correctly, and how many incorrectly. Such a matrix makes it simple to tell which classes are hard to discern from one another, as well as allows for deriving statistics such as users' accuracy, producers' accuracy, and overall accuracy \citep{foody1996fuzzyevaluation}, as well as variance if probability sampling was used. However, a standard confusion matrix is not applicable in the context of fuzzy classification, since misclassification in this case is not absolute, but rather a matter of degree \citep{foody2002accuracy}. Several solutions, such as distance \citep{foody1996fuzzyevaluation}, cross-entropy and mutual information \citep{lu2007methods} have been suggested as alternatives for fuzzy classification accuracy assessment.

%Visualisation of the fuzzy classification results is also challenging, since each class effectively is a single-channel raster of its own. Three such classes can easily be combined into RGB channels for visualisation, but with more classes it is no longer possible. There have been attempts to develop a method based on the hue, saturation and intensity colour model to allow for a larger number of distinct classes to be visualised on a single raster \citep{hengl2004fuzzycmeans}. There is also a possibility of ``hardening'' the classification when high accuracy visualisation is not needed, and making use of multiple RGB rasters side-by-side when it is needed.

%All in all, even though global land cover mapping has been a focus of many other studies, existing global land cover maps still have room for improvement. This can be achieved by using data from the sensors of newer satellites such as PROBA-V, as well as performing fuzzy classification as opposed to the traditional hard classification.

%An important long-term goal of land cover mapping is to have a highly accurate fuzzy land cover classification method at a global scale. Since there are many different methods that could be used to achieve such a goal, and since the amount of data generated by satellites covering the entire world is enormous, this thesis focuses on comparing several different fuzzy classification methods on a subset of global imagery provided by the PROBA-V satellite.

%PROBA-V is a relatively new satellite, with few studies using it for land cover classification so far. A number of studies have used simulated PROBA-V data in preparation for its launch \citep{stathakis2014probavurban,roumenina2013probavcrops,bartalev2014probavboreal}. After launch, most studies have focused on its potential for crop classification \citep{roumenina2015probavcrops,durgun2016crop,lambert2016cropland}. All of these studies have only used the traditional hard classification. As such, the potential of using PROBA-V data for fuzzy classification in order to obtain a more accurate representation of land cover at global scale has never been tested before.

%Another problem of land cover classification is that there is high disagreement between different land cover products on mixed classes, such as mixed forests \citep{Herold2008lccomparison}. Mixed forests and wetlands are challenging to detect and classify using hard classification and optical remote sensing data, so it is important to check how well large-scale fuzzy classification performs on these classes. Therefore this thesis focused specifically on the region of ecotone from boreal forests to temperate broadleaf forests in Europe. This region also includes a large number of boreal wetlands.

%In addition, there is a wide variety of methods that can be used for fuzzy classification. Most studies on this topic focus on comparing only two methods at once. In this thesis, four of them have been tested: fuzzy \textit{c}-means, neural networks, random forest regression and multiclass gradient boosting. All of these methods are based on fundamentally different approaches to classification, so the results give insight into the possible performance of similar methods as well. In addition, the last of the mentioned methods has not yet been tested on fuzzy classification itself (although related methods have been).

%When it comes to global image processing, one problem is the amount of data and the time it takes to process it. Some algorithms are known to suffer from the curse of dimensionality, where a linear increase in input data results in a geometric increase in processing time \citep{walton2008subpixelrf}. Therefore it is important to also measure the processing time of different classification methods.

@@

For deriving land cover fractions, we compared a wide array of machine learning regression methods: \ac{FNC}, \ac{MLP} \glspl{NN}, \ac{PLS} regression, Lasso regression, \ac{MLR}, \ac{RF} regression.
% * <Dainius Masiliūnas> 16:05:16 07 Aug 2019 UTC+0200:
% Perhaps also SVM and MESMA
In addition, as a baseline we also compare the results with an intercept model (all fractions always predicted to be equal).

@@

@@

Since we performed a manual covariate selection prior to running regularisation to remove highly correlated covariates, and having a further reduced set of covariates is useful for other algorithms, for this method we chose Lasso regression.
Lasso regression builds upon least squares, and thus the classes are predicted separately to each other, using the binary relevance approach as presented by \citet{karalas2016br}.
Therefore, the output fractions do not necessarily sum to 100\%.
As such, they were linearly scaled.
The lasso regression was performed using the \texttt{glmnet} package \citep{glmnet}.

@@

\ac{PLS} regression is a concept related to principal component regression and also aims at reducing the input covariates into latent variables that explain the most variance of the response variable in multidimensional space \citep{martens_multivariate_1992}.
As such, it is well suited for multicollinear data.
Due to this property, the \highlight[id=nt,comment={would you not agree for better comparability just use the same set of covariates?}]{full set of covariates} were used as input to this model, rather than the reduced (uncorrelated) set.
\ac{PLS} regression was performed using the package \texttt{pls} \citep{pls}.

@@

\Gls{FNC} regression, also called fuzzy nearest prototype, fuzzy $c$-means or fuzzy $k$-means, is a simple regression method where the pixel's membership of a class is determined by the distance of the pixel from the class centroid in feature space.
This method always produces output that sums to 100\%.
In this study, we used an implementation adapted from the package GSIF \citep{hengl2004fuzzycmeans}.

@@

\Glspl{NN} are a promising technique for land cover fraction mapping, as they allow both multiple inputs and multiple outputs, and, using the softmax activation function, also ensures that the result sums up to 100\%.
In this study, we used a multilayer perceptron with two layers of 256 neurons and softmax activation.
The \texttt{keras} package \citep{keras}, built upon TensorFlow, was used as the \ac{NN} implementation, and for the ability to make use of a graphics card to accelerate the neural network training process.

@@

\ac{RF} regression is a popular method for land cover classification that works by building a number of decision trees based on random subsets of the input training data, and averaging the ``votes'' of these individual trees.
\ac{RF} can only take one variable as the response variable, therefore there are two ways to obtain land cover fraction predictions from an \ac{RF} model.
One way is analogous to how we performed \ac{MLR}: giving labels of the dominant land cover class as the response variable, and then considering the proportion of votes from the trees as the proportions of land cover classes in the pixel.
The second way is binary relevance, where a separate model is trained to predict the presence of each class.
Then the predictions are scaled to sum up to 100\%.
In this study, we chose the second approach, so that the model for each class would be able to learn from the fractions of all the input data, rather than just the labels of the dominant land cover class.
The processing was done using the \texttt{ranger} package \citep{ranger}.

\ac{RF} was also used for assessing the covariate importance.
Since the number of covariates is imbalanced by group (i.e. there are merely 3 location covariates and 168 soil covariates, which are often collinear), covariate importance was assessed by permuting all covariates in a group and reporting the decrease in \ac{RMSE} resulting from that.

@@

In addition, we investigated applying histogram matching, by matching the histogram of the predictions back to the histogram of the training set.

@@

\begin{enumerate}
 \item Differences between all models, showing the RF has lowest RMSE
 \item Variable importance for RF; stressing importance of RS data by a model with only RS variables
 \item Spatial wall-to-wall demo map and spatial errors
\end{enumerate}

@@

\begin{table}
\centering
\begin{tabular}{llllll}
\toprule
\textbf{Model} & \textbf{\ac{RMSE} (\%)} & \textbf{\ac{MAE} (\%)} & \textbf{\acrshort{NSE}} & \textbf{\ac{OA} (\%)} & \textbf{Kappa} \\
% * <Dainius Masiliūnas> 14:04:01 13 Aug 2019 UTC+0200:
% Death to kappa
\midrule
Intercept
& 29.9  & 21.4  & 0     & 26±5  & 0.13±0.07 \\
\Gls{FNC}
& 24.4  & 13.5  & 0.33  & 53±4  & 0.42±0.06 \\
General linear model and \Gls{PLS}
& 21.6  & 12.7  & 0.48  & 56±4  & 0.43±0.05 \\
Lasso regression
& 21.4  & 12.6  & 0.49  & 56±4  & 0.43±0.05 \\
\Gls{MLR}
& 21.6  & 12.1  & 0.48  & 58±4  & 0.46±0.06 \\
Multivariate RF
& 18.7  & 10.0  & 0.61  & 65±3  & 0.55±0.05 \\
\Glspl{NN}
% * <Dainius Masiliūnas> 18:06:02 15 Apr 2020 UTC+0200:
% Double-check NN results - were they updated?
& 22.7  & 9.2   & 0.43  & 68±1  & 0.57±0.02 \\
\Gls{SVM} regression
& 20.7  & 8.9   & 0.52  & 69±2  & 0.58±0.03 \\
Cubist regression
& 18.1  & 8.1   & 0.63  & 72±2  & 0.63±0.03 \\
\Gls{RF} regression
& 17.3  & 9.4   & 0.66  & 67±4  & 0.57±0.05 \\
%\ensuremath{''} histogram matched
%& 22.0  & 9.7   & 0.43  & 66±2  & 0.58±0.03 \\
\ensuremath{''} two-step
& 19.9  & 8.2   & 0.56  & 71±2  & 0.62±0.02 \\
\ensuremath{''} three-step
& 19.4  & 8.4   & 0.58  & 71±3  & 0.62±0.04 \\
\ensuremath{''} median vote
& 20.7  & 7.9   & 0.52  & 71±1  & 0.63±0.02 \\
\ensuremath{''} \ensuremath{''} two-step
& 20.0  & 8.1   & 0.54  & 72±1  & 0.63±0.02 \\
\ensuremath{''} \ensuremath{''} three-step
& 20.2  & 7.9   & 0.54  & 72±2  & 0.64±0.02 \\
%\ensuremath{''} \ensuremath{''} histogram matched
%& 21.3  & 9.1   & 0.46  & 68±3  & 0.61      \\
\bottomrule
\end{tabular}
\caption{Accuracy statistics of the models tested.}
\label{tab-accuracy}
\end{table}

% Maps
\begin{figure}
    \includegraphics[width=\textwidth]{article-figures/maps/2019-03-26-rf-1m-bubble}
    \caption{Random Forest single model prediction residuals per class, spatially}
    \label{resid-rf-1m-uncor}
\end{figure}
\begin{figure}
    \includegraphics[width=\textwidth]{article-figures/maps/2019-03-26-rf-3m-bubble}
    \caption{Random Forest three-step model prediction residuals per class, spatially}
    \label{resid-rf-3m-uncor}
\end{figure}
\begin{figure}
    \includegraphics[width=\textwidth]{article-figures/maps/2019-04-10-rasterised-validation}
    \caption{Validation data spatially}
    \label{raster-validation}
\end{figure}
\begin{figure}
    \includegraphics[width=\textwidth]{article-figures/maps/2019-04-10-rasterised-rf-1m-uncor}
    \caption{Random Forest single model predictions spatially}
    \label{raster-rf-1m-uncor}
\end{figure}

@@

>>>>

% Minimalistic template for a generic article, geared towards MDPI guidelines
% Use PDFLaTeX
\documentclass[a4paper,10pt]{article}
% Packages provided by the MDPI template already
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{indentfirst}
\usepackage{booktabs} % For \toprule, etc.
\usepackage[sort&compress,sectionbib]{natbib}
\usepackage{graphicx}
\bibliographystyle{apalike}

@@

@@

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS-LC100}{CGLS-LC100}{Copernicus Global Land Services 100 m Land Cover}
\newacronym{CGLS}{CGLS}{Copernicus Global Land Services}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{IIASA}{IIASA}{International Institute for Applied Systems Analysis}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym{OSAVI}{OSAVI}{Optimised Soil-Adjusted Vegetation Index}
\newacronym{NDVI}{NDVI}{Normalised Difference Vegetation Index}
\newacronym{NDMI}{NDMI}{Normalised Difference Moisture Index}
\newacronym{NIRv}{NIRv}{\gls{NIR} of vegetation}
\newacronym{NIR}{NIR}{near infra-red}
\newacronym{SCM}{SCM}{subpixel confusion-uncertainty matrix}
\newacronym{FNC}{FNC}{fuzzy nearest centroid}
\newacronym{PLS}{PLS}{partial least squares}
\newacronym{MLR}{MLR}{multinomial logistic regression}
\newacronym{RF}{RF}{random forest}
\newacronym{TOC}{TOC}{Top-of-Canopy}
\newacronym{CART}{CART}{Classification and regression tree}
\newacronym{GLM}{GLM}{General linear regression model}
\makenoidxglossaries
% Disable abbreviating RF for now [text={random forest}]
%\glsunset{RF}

@@

\minisection{Highlights}

\begin{itemize}
    \item 11 statistical and machine learning models tested for land cover fraction mapping
    \item Random forest regression performed the best, with RMSE of 17.3 globally
    \item A three-step regression method proposed, which results in the lowest MAE of 7.9
    \item Remote sensing data was the most important for model predictions, but all auxiliary data also contributed significantly
    \item Global wall-to-wall land cover fraction maps produced and their errors assessed
\end{itemize}

@@

% \begin{enumerate}
%  \item Relevance of fractional LC mapping (link to user requirements and SDGs)
%  \item Studies that have looked into fractional LC mapping so far
%  \item Available fractional methods that were introduced in literature (wavelet transform etc.)
%  \item Methods that are tested in this paper
%  \item Global land cover maps
% \end{enumerate}

Land cover, as one of the key variables for monitoring a number of \glspl{SDG}, has received attention in recent years, and new global land cover maps have emerged to better map the current land cover of the world, as well as to track land cover change.
Some of the recent achievements have been the \ac{LC-CCI} map \citep{defourny2012cci} that aims at maintaining a long medium-resolution time series of global land cover for the climate community, \ac{CGLS-LC100} map that aims at a finer spatial resolution and high quality with yearly updates since 2015, and the \ac{FROM-GLC} map \citep{fromglc2019} that showcases the potential of land cover mapping at 10 m resolution.

@@

A potential solution to this issue is land cover fraction mapping, where instead of a discrete class, the proportion of every class in the legend is reported for every pixel of the map.
This is also called ``fuzzy'' or ``soft'' classification, and sometimes ``subpixel'' classification or ``linear mixture modelling'' \citep{Okeke2006fuzzyexponent}.
Related terms are ``super-resolution mapping'' and ``subpixel mapping'', but this concept relates also to the allocation of land cover fractions inside a pixel spatially, rather than just determining the fraction ratio per pixel.

Land cover fraction mapping has been attempted in the past, both in research \citep{adams_classification_1995, foody1996fuzzyevaluation, colditz_land_2011, sharma_assessing_2011, uma_shankar_wavelet-fuzzy_2011, dwivedi_optimisation_2012, lizarazo_quantitative_2012, gessner_estimating_2013, okujeni_generalizing_2018} and production (e.g. \citealp{hansen2000hardtree, hansen_continuous_2011, pengra_global_2015, hansen_global_2003, sexton_global_2013}).
Nevertheless, to date only \ac{CGLS-LC100} provides a global map with fractions of every major land cover class.

@@

@@

For deriving land cover fractions, we compared a wide array of machine learning regression methods: \ac{FNC}, \ac{GLM}, \ac{PLS} regression, Lasso regression, \ac{MLR}, \ac{MLP} \glspl{NN}, \glspl{SVM}, Cubist and \ac{RF} regression.
In addition, as a baseline we also compare the results with an intercept model (all fractions always predicted to be equal).

@@

\subsubsection{\Glsentryfull{GLM}}

\Gls{GLM}, also known as multivariate linear regression, is an extension to a regular linear regression that allows for multiple outcomes.
In this study, we used \gls{GLM} as a baseline model.
\Gls{GLM} does not normalise the output to add up to 100\%, and also allows negative predictions.
Therefore all predictions < 0 were capped to 0, and then all values were linearly rescaled per each prediction to add up to 100\%.

We also tested whether dropping any covariates would result in better predictions, using a sphericity test with Greenhouse-Geisser correction to compare models with different covariates.
However, only three covariates could be dropped and it did not improve the predictions, therefore we decided to keep all the covariates in the model.
\Gls{GLM} is implemented in base R \citep{r_2019}.

@@

We chose lasso regression, so that it would be possible to identify if there are any covariates that could be further omitted in addition to ones dropped in section \ref{sec-covariate-selection}.
The $\lambda$ parameter was chosen using cross-validation.

The lasso regression was performed using the \texttt{glmnet} package \citep{glmnet}.
This method is multivariate in the same way that \gls{GLM} is, and the output was clamped to positive numbers and rescaled to add up to 100\% the same way.

@@

\ac{PLS} regression is a concept related to principal component regression and also aims at reducing the input covariates into latent variables that explain the most variance of the response variable in multidimensional space \citep{martens_multivariate_1992}.
As such, it is well suited for multicollinear data.
This method is also multivariate in the same way that \gls{GLM} is, and the same method was used to clamp and rescale the predictions.
\ac{PLS} regression was performed using the package \texttt{pls} \citep{pls}.

@@

\Gls{FNC} regression, also called fuzzy nearest prototype, fuzzy $c$-means or fuzzy $k$-means, is a simple regression method where the pixel's membership of a class is determined by the distance of the pixel from the class centroid in feature space.
This method always produces output that sums to 100\%.

In this study, we used an implementation adapted from the package GSIF \citep{hengl2004fuzzycmeans}.
We tested both using a logistic regression to estimate the class centroids, as well as using a weighted average of the training data to do so, and ended up using the latter method as the result was more accurate.

@@

\Glspl{NN} are a promising technique for land cover fraction mapping, as they allow both multiple inputs and multiple outputs, and, using the softmax activation function, also ensures that the result sums up to 100\% with no need for any postprocessing.
In this study, we used a multilayer perceptron with three hidden layers with 128, 64 and 32 neurons per layer respectively.
Softmax activation was used to get the output.
The Nadam optimiser \citep{dozat_incorporating_2016} was used, with \gls{MAE} as the loss function.
The \texttt{keras} package \citep{keras}, built upon TensorFlow, was used as the \ac{NN} implementation, and for the ability to make use of a graphics card to accelerate the neural network training process.

@@

\ac{RF} regression is a popular method for land cover classification that works by building a number of decision trees based on random subsets of the input training data, and taking the mean or median of the ``votes'' of these individual trees \citep{breiman2001random}.
\ac{RF} can only take one variable as the response variable, therefore there are two ways to obtain land cover fraction predictions from an \ac{RF} model.
One way is analogous to how we performed \ac{MLR}: giving labels of the dominant land cover class as the response variable, and then considering the proportion of votes from the trees as the proportions of land cover classes in the pixel.
The second way is binary relevance, where a separate model is trained to predict the presence of each class \citet{karalas2016br}.
Then the predictions are scaled linearly to sum up to 100\%.
In the case of \gls{RF}, output values do not need to be clipped as their range is always within the constraints of the input data range.
In this study, we chose the binary relevance approach, so that the model for each class would be able to learn from the fractions of all the input data, rather than just the labels of the dominant land cover class.
The processing was done using the \texttt{ranger} package \citep{ranger}.

\ac{RF} was also used for assessing the covariate importance.
Permutation importance was used: each covariate was shuffled in turn, and the resulting decrease in \gls{RMSE} was calculated, comparing against the validation data set.
Since the number of covariates is imbalanced by group (i.e. there are merely 3 location covariates and 22 spectral covariates), and there may still be colinear covariates, we also tested calculating covariate importance by permuting all covariates in a group together.
However, the result was similar to the sum of permutation importance of each individual covariates in the group, which showed that the covariate selection process had eliminated all colinear covariates.
Thus we reported the more detailed covariate importance permuted per individual covariate.

\subsubsection{Cubist regression}

Cubist regression \citep{cubist} is based on \gls{RF} regression, with the major difference being that instead of using a threshold of a covariate for a split in the decision tree, Cubist instead uses a linear regression based on a subset of the data relevant for the split in question.
In addition, it features committees, a boosting technique that iteratively trains trees so as to learn from the previous ones.
For our dataset, 10 committees gave the best result.
Like with \gls{RF}, Cubist was used with the binary relevance method.
The \texttt{Cubist} package \citep{cubistpackage} was used for this.

\subsubsection{\Glsentryfull{SVM} regression}

\Glspl{SVM} are machine learning models that attempt to find the optimal boundary between the class clusters in feature space by constructing an optimal dividing hyperplane.
For land cover fraction classification, we used \gls{SVM} regression based on Least Squares \gls{SVM} \citep{suykens_least_1999}.
We used the binary relevance method, clamped the output values to the 0-100\% range and linearly scaled them to add up to 100\%.
The package \texttt{liquidSVM} \citep{liquidSVM} was suitable for this purpose, as it is optimised for processing large datasets such as ours.

\subsubsection{Other methods}

A few more methods were considered, but in the end turned out to be unworkable with our dataset.
Multivariate \gls{RF} in available in R provided by the \texttt{MutivariateRandomForest} package \citep{MultivariateRandomForest}, which would potentially allow better results than using regular \gls{RF} with binary relevance.
However, it is not optimised for large datasets, which did not allow to train it on more than 2000 samples at a time, making it incomparable to the rest of the tested methods.

Another model that was considered was SuperLearner \citep{SuperLearner}, an ensembling method that can combine multiple of the tested methods into one model, giving models different weights according to how well they are known to perform for a given task.
Here we ran into similar issues, as cross-validation is required for each model to train, which increases processing time dramatically.
Whereas setting the number of folds to a lower number resulted in worse results than when using a single classifier.

Lastly, multiple-endmember spectral unmixing using non-negative least squares was also attempted, but the results were significantly worse than even the intercept model, possibly because the class fractions are not a linear combination of the covariates.

@@

In addition, we investigated applying histogram matching, by matching the histogram of the predictions back to the histogram of the training set.
However, that turned out to decrease accuracy in all cases, as the cases where a model predicts a value close to the true value would be set to a more extreme value to compensate for other predictions that were not as accurate to begin with.
Therefore in the end histogram matching was not used.

@@

@@

\begin{table}
\centering
\begin{tabular}{llllll}
\toprule
\textbf{Model} & \textbf{\ac{RMSE} (\%)} & \textbf{\ac{MAE} (\%)} & \textbf{\acrshort{NSE}} & \textbf{\ac{OA} (\%)} & \textbf{Kappa} \\
\midrule
Intercept
& 29.9  & 21.4  & 0     & 26±5  & 0.13±0.07 \\
\Gls{FNC}
& 24.4  & 13.5  & 0.33  & 53±4  & 0.42±0.06 \\
\Gls{GLM}, \Gls{PLS}, Lasso
& 21.6  & 12.7  & 0.48  & 56±4  & 0.43±0.05 \\
%Lasso regression
%& 21.4  & 12.6  & 0.49  & 56±4  & 0.43±0.05 \\
\Gls{MLR}
& 21.6  & 12.1  & 0.48  & 58±4  & 0.46±0.06 \\
%Multivariate RF
%& 18.7  & 10.0  & 0.61  & 65±3  & 0.55±0.05 \\
\Glspl{NN}
& 22.7  & 9.2   & 0.43  & 68±1  & 0.57±0.02 \\
\Gls{SVM} regression
& 20.7  & 8.9   & 0.52  & 69±2  & 0.58±0.03 \\
Cubist regression
& 18.1  & 8.1   & 0.63  & 72±2  & 0.63±0.03 \\
\Gls{RF} regression
& \textbf{17.3}  & 9.4   & \textbf{0.66}  & 67±4  & 0.57±0.05 \\
%\ensuremath{''} histogram matched
%& 22.0  & 9.7   & 0.43  & 66±2  & 0.58±0.03 \\
\ensuremath{''} two-step
& 19.9  & 8.2   & 0.56  & 71±2  & 0.62±0.02 \\
\ensuremath{''} three-step
& 19.4  & 8.4   & 0.58  & 71±3  & 0.62±0.04 \\
\ensuremath{''} median vote
& 20.7  & 7.9   & 0.52  & 71±1  & 0.63±0.02 \\
\ensuremath{''} \ensuremath{''} two-step
& 20.0  & 8.1   & 0.54  & 72±1  & 0.63±0.02 \\
\ensuremath{''} \ensuremath{''} three-step
& 20.2  & \textbf{7.9}   & 0.54  & \textbf{72±2}  & \textbf{0.64±0.02} \\
%\ensuremath{''} \ensuremath{''} histogram matched
%& 21.3  & 9.1   & 0.46  & 68±3  & 0.61      \\
\bottomrule
\end{tabular}
\caption{Accuracy statistics of the models tested.}
\label{tab-accuracy}
\end{table}

@@

\Gls{RF} permutation importance was performed to test which variables were important for \gls{RF} regression predictions.
The single-step \gls{RF} model was selected for this as it achieved the lowest \gls{RMSE}.
The results are shown in figure \ref{fig-varimp}.

\begin{figure}
    \includegraphics[width=\textwidth]{article-figures/heatmaps/2020-05-12-varimp-heatmap-top10}
    \caption{Random Forest single model variable importance, top 10 covariates per category.}
    \label{fig-varimp}
\end{figure}
